Transactions Fraud Detection using Machine Learning and Nature Inspired Algorithms

### 2020-05-29 paper/main.tex
Update TODOs
Fix mistakes

% TODO: len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu
% TODO: mozeno spomenut oversampling/undersampling

@@

Since the dataset, if pretty large, firstly we \textit{undersampled} it to achieve the equal size of both classes. Then we shrunk it again to 5\% of its size, so we would be able to make as many experiments as possible (we did not have enough computing power). We did a train/test split of 80/20.

For each algorithm, we tried several population sizes (10, 20, and 50), for which we did 5 runs of the algorithm. We trained exactly 50 generations. For both, recursive feature selection and feature selection using nature-inspired algorithms we did 5 fold cross-validation. We did not experiment with algorithm parameters and we used their defaults which are described in the previous section.

@@

When comparing the population sizes, all of the algorithms achieved the best results when the population size was highest (50). The only exception is Cuckoo Search, which found the most optimal solution with population size 10, however the difference negligible. Table \ref{tab:nia_comparison}. shows the algorithm scores for each population. Figure \ref{fig:nia_train_score_by_algorithm}. shows the differences in scores between population size of 10 and 50.

@@

All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm}). These differences in results (means) between GWO and other algorithms are significant (t-test; $p-value=0.012$). Looking for the explanation, we found out that only Grey Wolf Algorithm and Bat Algorithm converged during the optimization run. Figure \ref{fig:nia_scores_by_run} shows the optimization of the Firefly Algorithm and Grey Wolf Optimizer, the difference is noticeable. If we fine-tuned the parameters of other algorithms, they might converge as well. The fact that GWO does not have any parameters (except the population size) was an advantage for it in our experiments.

@@

The best features, selected by the optimization process, were later used to train the final model on test data. These models were compared to a model trained on all features and the model trained on features selected by recursive feature elimination.

@@

% TODO: ake experiemnty sme spravili
% TODO: tabulka porovnania algoritmov
% TODO: learning along the way with consequences - napr undersampling/oversampling atd.

@@

% TODO: zhrnutie vysledkov, mozne vylepsenia a uskalia nasho pristupu

@@

>>>>

% TODO: Len kratko, co navrhujeme - feature selection na nejaky model pomocou nejakeho prirodou inspirovaneho algoritmu

@@

Since the dataset is pretty large, we firstly \textit{undersampled} it to achieve the equal size of both classes. Then we shrunk it again to 5\% of its size, so we would be able to make as many experiments as possible (since we did not have enough computing power). We did a train/test split of 80/20.

For each algorithm, we tried several population sizes (10, 20, and 50), for which we did 5 runs of each algorithm. We trained exactly 50 generations. For both, recursive feature selection and feature selection using nature-inspired algorithms we did 5 fold cross-validation. We did not experiment with algorithm parameters and we used their defaults, as we mentioned in the previous section.

@@

When comparing the population sizes, all of the algorithms achieved the best results when the population size was highest (50). The only exception is Cuckoo Search, which found the most optimal solution with population size of 10, however the difference negligible. Table \ref{tab:nia_comparison}. shows the algorithm scores for each population. Figure \ref{fig:nia_train_score_by_algorithm}. shows the differences in scores between population size of 10 and 50.

@@

All of the nature-inspired algorithms perform roughly the same, except the Grey Wolf Optimizer which outperformed the others (Figure \ref{fig:nia_train_score_by_algorithm}). These differences in results between GWO and other algorithms are significant (there is a sigificant difference in means, t-test; $p-value=0.012$). Looking for the explanation, we found out that only Grey Wolf Algorithm and Bat Algorithm converged during the optimization run. Figure \ref{fig:nia_scores_by_run} shows the optimization progress of the Firefly Algorithm and Grey Wolf Optimizer, Firefly Algorithm did not converge. If we fine-tuned the parameters of other algorithms, they might converge as well. The fact that GWO does not have any parameters (except the population size) might be an advantage for it in our experiments.

@@

The best features, selected in the optimization process, were later used to train the final model on test data. These models were compared to the model trained on all features and the model trained on features selected by recursive feature elimination.

@@

@@

@@

% TODO: concusion in the topic of transaction fraud detection
